Learning CNN filters from user-drawn image markers for coconut-tree image classification

### 2020-07-24 main.tex
Updates from Overleaf

The user selects a few \textcolor{red}{images from a training set} to compose a very small dataset $\D$ and draws labeled markers in image regions that best discriminate the classes. The convolution between an image $I$ and a filter $F\in {\cal F}$ \textcolor{red}{ enhances} regions whose local patterns are detected by $F$. We wish then to estimate the weights of $F$ such that those local patterns are characteristic of one given class. For sake of clarity, one must interpret the convolution operation at a pixel $p$ of $I$ as the inner product $\vect(P_I(p)) \cdot \vect(F)$, where $P_I(p)$ is a $k\times k$ patch with $m$ channels around $p$, $F$ is a filter with shape $k\times k \times m$, and $\vect$ is the vectorization operation. A filter $F$ can discriminate a class $i \in \{1,2,\ldots,c\}$ among $c$ classes when it generates positive results for local patterns $\vect(P_I(p))$ of class $i$ and negative results for patterns from other classes. That is, $\vect(F)$ is the normal vector of a hyperplane in $\R^{k\times k\times m}$, which detects the patterns of class $i$ at its positive side. We wish to estimate ${\cal F}$ such that its filters will detect patterns from all classes in different positions $p$.

For a problem with $c$ classes where $\lambda(p)=i\in \{1,2,\ldots,c\}$ is the label of a marker pixel $p$ from class $i$, let ${\cal M}_I$ be a set of marker pixels drawn in image $I\in {\cal D}$ and $P_I(p)$ be the respective patch around a pixel $p\in {\cal M}_I$. Let ${\cal P}_i$ be the set of all patches around marker pixels from all images $I\in {\cal D}$, with representative examples  $\vect(P_I(p))$ from class $i\in \{1,2,\ldots,c\}$. 
\begin{eqnarray}
{\cal P}_i & = & \bigcup_{I\in {\cal D},p\in {\cal M}_I, \lambda(p) = i}{P_I(p)}.
\label{eq:pset}
\end{eqnarray}
A clustering operation on each set ${\cal P}_i$, $i\in [1,c]$, guarantees groups with similar local patterns for each class. The groups must be shifted to the origin of $\R^{k\times k\times m}$ by subtracting the mean value of all patches, set $\mathcal{P} = \bigcup_{i \in \{1, 2, \ldots, c\}}{\mathcal{P}_i}$. Additionally, the standard deviation of all patches is computed and used for standardization. This operation allows batch normalization of image sets using the mean and standard deviation of the patches from the image markers -- i.e.,  \emph{a marker-based batch normalization}. The centroid of each cluster defines the weights of each filter $F\in {\cal F}$, and we force norm $\|\vect(F)\|=1$ to avoid preferences among them. The centralization for filter definition is paramount to eliminate activations in regions from other classes, whose patterns fall in the negative side of the hyperplane with normal $\vect(F)$. Figure~\ref{fig:filter} illustrates three groups with color-coded samples from two classes in a hypothetical 2D feature space. The marker-based batch normalization and convolution with filters obtained as centroids of the three groups in Figure~\ref{fig:filter}a creates a new 3D feature space (Figure~\ref{fig:filter}b), in which the classes can be more easily detected by two filters in distinct activation bands.

@@

We use a dataset from~\cite{8899005}, which contains regions of aerial images with and without coconut trees from the Kingdom of Tonga, as acquired in October 2017~\footnote{The images are available in \url{blog.werobotics.org/2018/01/11/open-ai-challenge-2}.}. Each region is classified as containing one or none coconut tree, but part of coconut trees from adjacent regions might appear near the region's border. The Humanitarian OpenStreetMap community annotated the regions. The dataset consists of 13587 regions with $90 \times 90$ pixels \textcolor{red}{with spatial resolution of $8$\si{\centi\metre}}, being 10268 and 3319 regions annotated as containing and not containing coconut trees, respectively. \textcolor{red}{The region images contain three channels: red, green, and blue. We convert them to the CIE L*a*b* color space and normalize each channel within $[0,1]$}.

@@

\textcolor{red}{First, we evaluated H1 and H2 by randomly selecting 200 images for the training set, 1000 images for a validation set, and the remaining 12387 images were left in the testing set. The validation set was used to choose empirically the architecture of the network shown in Figure~\ref{fig:arch}. The convolutional layer has $60$ filters with dimension $7 \times 7$ and 3 color channels. It includes marker-based normalization before convolution, ReLu activation, and a max-pooling operation with a window of dimension $3 \times 3$. Since it is the single and last convolutional layer, we applied a stride of $4$ in max-pooling and batch normalization to create an input as close as possible to the one of the \textcolor{red}{MLP} classifier used in the popular VGG-16~\cite{simonyan2014very}. We then vectorized and projected the training images by t-SNE~\cite{maaten2008visualizing} (Figure~\ref{fig:pipeline-xtractor}). Four representative images, only two per class, were chosen by the user to draw markers (Figure~\ref{fig:markers}). These images must come from regions in the projection populated by samples from the same class (either red or green points). Those regions indicate groups of images from the same class in the input feature space. The markers must be placed in regions that best distinguish the coconut and non-coconut classes.}

@@

\section*{Acknowledgements}

>>>>

The user selects a few \textcolor{red}{images from a training set} to compose a very small dataset $\D$ and draws labeled markers in image regions that best discriminate the classes. The convolution between an image $I$ and a filter $F\in {\cal F}$ \textcolor{red}{ enhances} regions whose local patterns are detected by $F$. We wish then to estimate the weights of $F$ such that those local patterns are characteristic of one given class. For sake of clarity, one must interpret the convolution operation at a pixel $p$ of $I$ as the inner product $\vect(P_I(p)) \cdot \vect(F)$, where $P_I(p)$ is a $k\times k$ patch with $m$ \textcolor{red}{bands} around $p$, $F$ is a filter with shape $k\times k \times m$, and $\vect$ is the vectorization operation. A filter $F$ can discriminate a class $i \in \{1,2,\ldots,c\}$ among $c$ classes when it generates positive results for local patterns $\vect(P_I(p))$ of class $i$ and negative results for patterns from other classes. That is, $\vect(F)$ is the normal vector of a hyperplane in $\R^{k\times k\times m}$, which detects the patterns of class $i$ at its positive side. We wish to estimate ${\cal F}$ such that its filters will detect patterns from all classes in different positions $p$.

For a problem with $c$ classes where $\lambda(p)=i\in \{1,2,\ldots,c\}$ is the label of a marker pixel $p$ from class $i$, let ${\cal M}_I$ be a set of marker pixels drawn in image $I\in {\cal D}$ and $P_I(p)$ be the respective patch around a pixel $p\in {\cal M}_I$. Let ${\cal P}_i$ be the set of all patches around marker pixels from all images $I\in {\cal D}$, with representative examples  $\vect(P_I(p))$ from class $i\in \{1,2,\ldots,c\}$. 
\begin{eqnarray}
{\cal P}_i & = & \bigcup_{I\in {\cal D},p\in {\cal M}_I, \lambda(p) = i}{P_I(p)}.
\label{eq:pset}
\end{eqnarray}
A clustering operation on each set ${\cal P}_i$, $i\in [1,c]$, guarantees groups with similar local patterns for each class. The groups must be shifted to the origin of $\R^{k\times k\times m}$ by subtracting the mean value of all patches, set $\mathcal{P} = \bigcup_{i \in \{1, 2, \ldots, c\}}{\mathcal{P}_i}$. Additionally, the standard deviation of all patches is computed and used for standardization. This operation allows batch normalization of image sets using the mean and standard deviation of the patches from the image markers -- i.e.,  \emph{a marker-based batch normalization}. The centroid of each cluster defines the weights of each filter $F\in {\cal F}$, and we force norm $\|\vect(F)\|=1$ to avoid preferences among them. The centralization for filter definition is paramount to eliminate activations in regions from other classes, whose patterns fall in the negative side of the hyperplane with normal $\vect(F)$. Figure~\ref{fig:filter} illustrates three groups with color-coded samples from two classes in a hypothetical 2D feature space. The marker-based batch normalization and convolution with filters obtained as centroids of the three groups in Figure~\ref{fig:filter}a creates a new 3D feature space (Figure~\ref{fig:filter}b), in which the classes can be more easily detected by two filters in distinct activation \textcolor{red}{bands}.

@@

We use a dataset from~\cite{8899005}, which contains regions of aerial images with and without coconut trees from the Kingdom of Tonga, as acquired in October 2017~\footnote{The images are available in \url{blog.werobotics.org/2018/01/11/open-ai-challenge-2}.}. Each region is classified as containing one or none coconut tree, but part of coconut trees from adjacent regions might appear near the region's border. The Humanitarian OpenStreetMap community annotated the regions. The dataset consists of 13587 regions with $90 \times 90$ pixels \textcolor{red}{with spatial resolution of $8$\si{\centi\metre}}, being 10268 and 3319 regions annotated as containing and not containing coconut trees, respectively. \textcolor{red}{The region images contain three bands: red, green, and blue. We convert them to the CIE L*a*b* color space and normalize each band within $[0,1]$}.

@@

\textcolor{red}{First, we evaluated H1 and H2 by randomly selecting 200 images for the training set, 1000 images for a validation set, and the remaining 12387 images were left in the testing set. The validation set was used to choose empirically the architecture of the network shown in Figure~\ref{fig:arch}. The convolutional layer has $60$ filters with dimension $7 \times 7$ and 3 color \textcolor{red}{bands}. It includes marker-based normalization before convolution, ReLu activation, and a max-pooling operation with a window of dimension $3 \times 3$. Since it is the single and last convolutional layer, we applied a stride of $4$ in max-pooling and batch normalization to create an input as close as possible to the one of the \textcolor{red}{MLP} classifier used in the popular VGG-16~\cite{simonyan2014very}. We then vectorized and projected the training images by t-SNE~\cite{maaten2008visualizing} (Figure~\ref{fig:pipeline-xtractor}). Four representative images, only two per class, were chosen by the user to draw markers (Figure~\ref{fig:markers}). These images must come from regions in the projection populated by samples from the same class (either red or green points). Those regions indicate groups of images from the same class in the input feature space. The markers must be placed in regions that best distinguish the coconut and non-coconut classes.}

@@

\section*{Acknowledgments}
